Supplement 2

### 2020-11-17 paper/paper.Rmd
edits

Missing data are values that should have been observed, but were not. Since the values are unobserved, this can mean the population of interest might not be properly sampled, inducing bias into a study's results. Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results [@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go missing many ways. For example, accidentally skipping a survey question, equipment failure, or intentionally not recording values. Before analysis, missing data must be handled, with values typically deleted or imputed. There is no universal approach to handling missing data. Contextual factors---such as, the study design and objective, and pattern of missingness---determine how missings should be handled, on a case-by-case basis [@Sainani2015].

@@

Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), `r binom_ci_data_shared$mean_pct` (95% CI: `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration---especially in contrast
to many of the papers that included (potentially meaningless) tests of
normality (e.g., Kruskall-Wallis test).

@@

Broadly, there are two types of missing values in data: implicit and explicit missings. Explicit values are missing, but recorded; whereas information in the data. For example, in \@ref(fig:implicit-missings), player Koenen has missing values for quarters two and four. Sometimes values like these can be logically imputed, as it might be known that these values are recorded as 0 in this format, rather than NA. Other places missing data can arise include: through joins when merging data without corresponding values, surveys, an inability to collect a biological sample (e.g., venous blood), equipment malfunction, failure, or not being worn.

@@

We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017]). The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in a simulated injury dataset. To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].

@@

Missing values can occur in the both the predictors (the "independent variables") and the response variable (the "dependent variables"). Care should be taken when imputing the outcome variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis. Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the outcome as they describe the uncertainty---see Chapter 15 in [@McElreath2020]. Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].

>>>>

Missing data are values that should have been observed, but were not. Since the values are unobserved, this can mean the population of interest might not be properly sampled, inducing bias into a study's results. Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results [@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go missing in many ways. For example, accidentally skipping a survey question, equipment failure, or intentionally not recording values. Before analysis, missing data must be handled, with values typically deleted or imputed. There is no universal approach to handling missing data. Contextual factors---such as, the study design and objective, and pattern of missingness---determine how missings should be handled, on a case-by-case basis [@Sainani2015].

@@

Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), `r binom_ci_data_shared$mean_pct` (95% CI: `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration.

@@

Broadly, there are two types of missing values in data: implicit and explicit missings. Explicit values are missing, but recorded; whereas information in the data. For example, in Figure 1, player Koenen has missing values for quarters two and four. Sometimes values like these can be logically imputed, as it might be known that these values are recorded as 0 in this format, rather than NA. Other places missing data can arise include: through joins when merging data without corresponding values, surveys, an inability to collect a biological sample (e.g., venous blood), equipment malfunction, failure, or not being worn.

@@

We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017]). The panels in Figure 2 give an overall sense of the extent of missing and complete data, in a simulated injury dataset. To learn more about exploring missing values, we recommend the vignettes in *naniar* [@naniar], and the methods in [@Tierney2018].

@@

Missing values can occur in the both the predictors (the "independent variables") and the response variable (the "dependent variables"). Care should be taken when imputing the response variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis. Methods that provide multiple imputed values for a response variable, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the response variable as they describe the uncertainty---see Chapter 15 in [@McElreath2020]. Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].

### 2020-11-17 paper/supplement-1.Rmd
update figures and try to fix tables
edit writing and formatting.
edits

---
title: Supplement 1
output: html_document
bibliography: supp1.bib
---

# Missing data examples relevant to football

One way to describe patterns of missing values is through Rubin's [@Rubin1976]
three missing data mechanisms: Missing Completely at Random (MCAR), Missing at
Random (MAR), and Missing Not at Random (MNAR). These mechanisms help describe
whether the missingness occurs by chance (MCAR), due to some observed variable
(MAR) or an unobserved variable (MNAR), with MCAR being the least harmful, and
MAR and MNAR biased. Here, we provide an example of data MAR, MCAR and MNAR,
relevant to football.

>>>>

---
title: Supplement 1
output: pdf_document
bibliography: supp1.bib
---

# Missing Data Examples Relevant to Australian Football

One way to describe patterns of missing values is through Rubin's [@Rubin1976]
three missing data mechanisms: Missing Completely at Random (MCAR), Missing at
Random (MAR), and Missing Not at Random (MNAR). These mechanisms help describe
whether the missingness occurs by chance (MCAR), due to some observed variable
(MAR) or an unobserved variable (MNAR), with MCAR being the least harmful, and
MAR and MNAR biased. Here, we provide an example of data MAR, MCAR and MNAR,
relevant to Australian Football.

### 2020-11-17 paper/supplement-2.Rmd
edit paper
edits

```{r check}
loadd(n_screened)
print(n_screened)
```

@@

```{r search-table}
table_terms <- tibble(
  SPORTDiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
    "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football"),
  Embase = c("(exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR 'performance analysis' OR (('performance'/exp OR performance) AND ('analysis'/exp OR analysis)) OR 'performance'/exp OR performance OR 'sport psychology'/exp OR 'sport psychology' OR (('sport'/exp OR sport) AND ('psychology'/exp OR psychology)) OR 'coaching'/exp OR coaching) AND [2019-2019]/py",
              "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py"),
  Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
              "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football")

@@

```{r agreement}
md_reported <- agree_var(gs_football, any_missing_data_statement) %>% 
  percent(scale = 1,
          accuracy = 0.1)
```

Articles were surveyed to determine: (a) whether a missing data
statement was included, or if missing data were acknowledged in a
figure/table; where data were missing, whether (b) data were considered
MCAR, MAR or MNAR; (c) imputation was used; (d) the method used for imputation; 
and (e) whether articles that did not include a missing data
statement shared their data. Two authors independently extracted the
data. Agreement was `r md_reported`; discrepancies were resolved by DNB. Findings are
reported as the proportion and 95% confidence interval (CI). Confidence
intervals were calculated using the Clopper-Pearson method for the
binomial distribution [@Clopper1934], via the 'binom' package [@binom] in R [@rcore].

@@

The search returned a total of 3,863 articles. Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure \@ref(fig:prisma)). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data, `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness. While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration--especially in contrast to many of the papers that included (potentially meaningless) tests of normality (e.g., Kruskall-Wallis test).

There were `r nrow(gs_football_missings_stated)` studies that provided some statement about missing data. A breakdown of the extracted information from these studies is provided in Table \@ref(tab:extracted-info).

@@

>>>>

```{r check}
loadd(n_screened)
```

@@

```{r search-table}
table_terms <- tibble(
  SPORTDiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
                  "AND",
    "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football"),
  Embase = c("(exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR 'performance analysis' OR (('performance'/exp OR performance) AND ('analysis'/exp OR analysis)) OR 'performance'/exp OR performance OR 'sport psychology'/exp OR 'sport psychology' OR (('sport'/exp OR sport) AND ('psychology'/exp OR psychology)) OR 'coaching'/exp OR coaching) AND [2019-2019]/py",
             "AND",
              "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py"),
  Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
             "AND",
              "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football")

@@

Articles were surveyed to determine: (a) whether a missing data
statement was included, or if missing data were acknowledged in a
figure/table; where data were missing, whether (b) data were considered
MCAR, MAR or MNAR; (c) imputation was used; (d) the method used for imputation; 
and (e) whether articles that did not include a missing data
statement shared their data. Two authors independently extracted the
data. Discrepancies were resolved by DNB. Findings are
reported as the proportion and 95% confidence interval (CI). Confidence
intervals were calculated using the Clopper-Pearson method for the
binomial distribution [@Clopper1934] via the *binom* package [@binom] in R [@rcore].

@@

@@

The search returned a total of 3,863 articles. Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure \@ref(fig:prisma)). The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`). Of the articles that did not mention missing data, `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness. While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration.

There were `r nrow(gs_football_missings_stated)` studies that provided a statement about missing data. A breakdown of the extracted information from these studies is provided in Table \@ref(tab:extracted-info).
